Ontario Graduate Scholarship: Plan of Study

### 2021-04-09 latex/ogs/main.tex
OGS 2021 update

When a gymnast is performing a swing-up routine on a bar, they do not kick their
feet at specified time intervals. Rather, a more natural human motion is to
smoothly swing their legs based on their current speed and position
\cite{pendulum_length_giant_gymnastics}. In doing so, they are able to generate
enough momentum to perform backflips on the bar. When controlling mechanical
systems, standard practice is to have a robot's actuators track a function of
time. Much like the gymnast, it can be more natural to have the robot track a
function of speed and position instead. The recent techniques of virtual
constraints offer this ability to produce realistic, biologically plausible
motion.

Mechanical systems can be constrained in two ways. Holonomic constraints
restrict position (for instance, car tires remain a fixed distance apart), while
nonholonomic constraints restrict velocities (such as how car tires cannot
slide sideways). Previous research has shown that one can enforce holonomic
constraints through a robot's actuators, using the method of "virtual holonomic
constraints" (VHCs) \cite{vhcs_for_el_systems}. VHCs have been used to control
walking robots \cite{stable-walking}, autonomous bicycles \cite{bicycle},
gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
among others. These VHCs are a special case of the more general virtual
nonholonomic constraints (VNHCs). In my master's thesis \cite{my-thesis} I
developed a theory for enforcing VNHCs and designed one which injects energy
into the gymnastics robot created by Wang \cite{xingbo-thesis}, thereby enabling
it to perform backflips on a bar.

@@

Mohammadi \textit{et.al.} \cite{manfredi-orbit-stabilization} showed that a
single VHC can result in several motion patterns (known as ``orbits"), from
which one specific motion pattern can be chosen if the robot is initialized with a
configuration that is ``close enough" to it. However, their method does not
work for all starting configurations of the robot. We propose to extend this
work by investigating a method for transitioning between VNHCs.

VNHCs have been studied with applications to human-robot cooperation
\cite{vnhc-human-robot-coop} and walking robots \cite{vnhc-biped-robot}. Horn
\textit{et.al.} \cite{hybrid_zero_dynamics_bipedal_nhvcs} have also examined
properties of VNHCs for general mechanical systems and applied their results to
stabilizing a chosen gait on walking robots. Our approach is novel since we 
assume one can stabilize two distinct VNHCs independently, and will develop
techniques to transition between them while maintaining safety constraints.
Doing so will allow one to chain together VNHCs which drive a robot's
configuration towards a desired orbit, regardless of the initial configuration.

First we will study mathematical conditions under which one can transition
between two VNHCs. These conditions will allow us to create a control algorithm
to switch between motion patterns. After completing this mathematical
formulation, we will test the theory on a gymnastics robot by designing a
gymnastics routine which is performed solely by transitioning between several different
VNHCS.

The advancements in mathematics from this research will bring improvements to
the control of other autonomous systems. The ability to transition between
complex constraints will allow for more expressive motion and may become
a standard technique for controlling biologically inspired robots.

>>>>

When a gymnast on a horizontal bar wants to swing up to do backflips, they
typically perform a technique called a ``giant".
This technique involves the gymnast swinging their legs based on their current
speed and position, thereby generating enough momentum to rotate around the bar
\cite{pendulum_length_giant_gymnastics}.

Standard practice in robotics begins with finding a trajectory over time
replicating a desired behaviour, then using the robot's actuators to track this
time-based trajectory. 
While this often works for simple systems, it can cause timing delays, it is not
robust to external disturbances, and it is not what humans do.
A more natural approach is to constrain the robot's actuators according to
some function of speed and position.
This type of behaviour is well described by the method of virtual constraints.

There are two types of constraints on a mechanical system:
\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
but not fly),
while \textit{nonholonomic} constraints restrict position and
velocity (e.g. cars cannot slide sideways).
One can design a specific constraint for a robot and enforce it using the
actuators.
Since this constraint is not enforced by physics, it is known as a ``virtual"
constraint
\cite{vhcs_for_el_systems}.
Virtual holonomic constraints (VHCs) have been used to control walking robots
\cite{stable-walking}, autonomous bicycles \cite{bicycle},
gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}.
Virtual nonholonomic constraints (VNHCs) have been used for human-robot
cooperation \cite{vnhc-human-robot-coop} and for improving robustness of walking
robots \cite{vnhc-biped-robot,hybrid_zero_dynamics_bipedal_nhvcs}, among other
applications.

@@

Virtual constraints have proven useful for generating biologically-inspired
behaviour in robotic systems.
In my master's thesis, I demonstrated that VNHCs can also inject energy into
robots:
I designed a VNHC which enabled a gymnastics robot to gain energy
and perform backflips on a horizontal bar \cite{my-thesis}.
Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
safely increase their momentum.
Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
safely slow down in a realistic and reliable manner.

We propose to study the energy injection and dissipation properties of VNHCs.
First, we will find mathematical conditions under which VNHCs are guaranteed to
inject/dissipate energy in abstract mechanical systems.
Then we will design a means of safely transitioning between two different VNHCs.
Finally, we will test the theory on a gymnastics robot by designing a gymnastics
routine which is performed solely by transitioning between different virtual
constraints.

The advancements in mathematics from this research will bring improvements to
the control of autonomous systems. 
The ability to safely transition between complex constraints by adding or
removing energy will allow for more expressive motion, and may become a standard
technique for controlling biologically inspired robots.
